Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision

### 2024-02-19 Milestone_3/report/report.Rmd
update milestone 3

---
title: "Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision"
author:
- familyname: Li
  othernames: Weihao (Patrick)
  address: Monash University
  qualifications: PhD student
email: weihao.li@monash.com
phone: (04) 0459 1219
department: Department of\newline Econometrics &\newline Business Statistics
organization: "Progress review" 
bibliography: report.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  monash::report:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
library(visage)
library(tidyverse)
```

@@

When conducting model diagnostics, despite the availability of numeric summaries endorsed by finite or asymptotic properties, graphical representations of data are often preferred or required by data analysts. The preference for visual diagnostics is attributed to its intuitive nature and the possibility of discovering unexpected abstract and unquantifiable insights. In the context of regression diagnostics, a common practice is to plot residuals against fitted values, which serves as a starting point for evaluating the adequacy of the fit and verifying the underlying assumptions. Other visualization techniques such as histograms, Q-Q plots and box plots can be used to identify potential issues with assumptions made about the data, such as linearity, normality, or homoscedasticity.

@@

The reliance of human assessment is a fundamental aspect of visual tests, but it may restrict its widespread usage. The lineup protocol is unsuitable for large-scale applications, due to its high labor costs and time requirements. Moreover, it presents significant usability issues for individuals with visual impairments, resulting in reduced accessibility.

Modern computer vision models offer a promising solution to this challenge. As a subfield of AI, computer vision with its modern deep learning architectures has successfully resolved numerous critical problems in automation. The development of the convolutional neural network (CNN) by @fukushima_neocognitron_1982 was inspired by the vision processing in living organisms. The modern computer vision model typically utilizes deep neural networks with convolutional layers, which leverage the hierarchical pattern in data and provide regularized versions of fully-connected layers. This approach downscales and transforms images by summarizing information in a small space. Numerous studies have shown that it can effectively tackle vision tasks, such as image recognition [@rawat_deep_2017], computer-aided diagnosis [@lee_image_2015], pedestrian detection [@brunetti_computer_2018], and facial recognition [@emami_facial_2012].

Utilizing computer vision models in reading data plot is not a common choice. Nevertheless, certain fields have adopted this idea by applying computer vision models in reading recurrence plots for time series regression [@ojeda_multivariate_2020], time series classification [@chu_automatic_2019; @hailesilassie_financial_2019; @hatami_classification_2018; @zhang_encoding_2020], anomaly detection [@chen_convolutional_2020], and pairwise causality analysis [@singh_deep_2017]. However, the assessment of lineups with computer vision models is a relatively novel area of study.

@@

In order to examine the potential applicability of integrating visual inference techniques into regression diagnostics, a pilot study was conducted in the first year with the participation of 64 individuals, which was followed by a formal study involving 443 participants in the subsequent year. The participants were presented with lineups consisting of 20 plots, where a residual plot was embedded along with 19 null plots, drawn with data simulated using the residual rotation technique. The study considered two primary forms of residual departures in multiple linear regression, namely non-linearity and heteroskedasticity. To enrich the visual features, different fitted value distributions, including normal, uniform, lognormal, and discrete, were also incorporated.

As shown in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower}, the study revealed that appropriate conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions compared to visual tests evaluated by humans, leading to excessive rejections even when downstream analysis and outcomes would not be significantly impacted by the small departures from a good fit. Specifically, conventional tests conclude that the model fit has issues nearly twice as frequently as humans would, and they frequently reject departures in the form of non-linearity and heteroskedasticity that are not visible to humans. For instance, based on the row of scatter plots at the bottom of Figure \ref{fig:polypower}, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5. The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25.

![Comparison of power between different tests for detecting non-linearity. The power curves are estimated using logistic regression, and the horizontal lines of dots represent non-reject and reject results from human observers for each lineup. The visual test has multiple power curves estimated from bootstrap samples. The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.\label{fig:polypower}](polypower-1.png)

![Comparison of power between different tests for detecting heteroskedasticity. Main plot shows the power curves, with dots indicating non-reject and reject in visual testing of lineups. The multiple lines for the visual test arise from estimating the power on many bootstrap samples. The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.\label{fig:heterpower}](heterpower-1.png)

These findings emphasize the crucial role of graphical diagnostics and support the integration of visual inference into regression diagnostics. Moreover, they provide a compelling rationale for the development of an automated visual inference system to evaluate lineups of residual plots. The comprehensive details of the research methodology, along with other intriguing findings can be found in the attached paper.

@@

2. Chapter Two presents a comparative study between visual testing and conventional testing in the context of regression diagnostics. This chapter will be identical to the attached paper.

3. Chapter Three discusses the development of an automatic visual inference system for evaluating lineups of residual plots. The computer vision models used in this chapter are primarily trained on simulated data from linear models that violate classical assumptions such as linearity and homoscedasticity. The chapter concludes with an evaluation of the performance of the system, comparing the power and characteristics of human subject-assessed visual tests with computer-assessed visual tests.

4. Chapter Four provides detailed description of the open-source software developed during the course of the research. This includes the R package `bandicoot` for code maintenance with object-oriented programming, `visage` for running visual inference experiments on linear regression models, and one or multiple R packages for employment of the automatic visual inference system.  Additionally, an online application of the automatic visual inference system is introduced. This chapter may be splitted into multiple chapters to provide a more in-depth description of each tool.

@@

The timetable of my PhD study from April 2023 to August 2024 is provided in Table \ref{tab:timetable}.

```{r}
data.frame(Date = c("April", 
                    "May",
                    "June",
                    "July",
                    "Sep",
                    "Oct",
                    "Nov",
                    "Dec",
                    "Mar",
                    "Aug"), 
           Event = c("Submit the abstract of the first paper to Australian Statistical Conference (ASC)", 
                     "Submit the first paper",
                     "Submit a poster for the first paper to Institute of Electrical and Electronics Engineers (IEEE) Visualization Conference",
                     "Leave for a month",
                     "Finalize the computer vision model",
                     "Attend the IEEE Visualization Conference",
                     "Start the web interface development",
                     "Attend the ASC",
                     "Submit the second paper",
                     "Submit thesis")) %>%
  kableExtra::kable(booktabs = TRUE,
                    label = "timetable",
                    caption = "Timetable till thesis submission") %>%
  kableExtra::kable_styling(latex_options = "scale_down") %>%
  kableExtra::group_rows("2023", 1, 8) %>%
  kableExtra::group_rows("2024", 9, 10)
```

>>>>

---
title: "Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision"
author:
- familyname: Li
  othernames: Weihao (Patrick)
  address: Monash University
  qualifications: PhD student
email: weihao.li@monash.com
phone: (04) 0459 1219
department: Department of\newline Econometrics &\newline Business Statistics
organization: "Pre-submission review" 
bibliography: report.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  monash::report:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=FALSE, messages=FALSE, warning=FALSE)
library(visage)
library(tidyverse)
```

@@

When conducting model diagnostics, graphical representations of data are often preferred or required by data analysts. The preference for visual diagnostics is attributed to its intuitive nature and the possibility of discovering unexpected abstract and unquantifiable insights. In the context of regression diagnostics, a common practice is to plot residuals against fitted values, which serves as a starting point for evaluating the adequacy of the fit and verifying the underlying assumptions. Other visualization techniques such as histograms, Q-Q plots and box plots can be used to identify potential issues with assumptions made about the data, such as linearity, normality, or homoscedasticity.

@@

The reliance of human assessment is a fundamental aspect of visual tests, but it may restrict its widespread usage. The lineup protocol is unsuitable for large-scale applications, due to its high labour costs and time requirements. Moreover, it presents significant usability issues for individuals with visual impairments, resulting in reduced accessibility.

Modern computer vision models are well-suited for addressing this challenge. They rely on deep neural networks with convolutional layers [@fukushima_neocognitron_1982]. These layers leverage hierarchical patterns in data, downsizing and transforming images by summarizing information in a small space. Numerous studies have demonstrated the efficacy of convolutional layers in addressing various vision tasks, including image recognition [@rawat_deep_2017]. Despite the widespread use of computer vision models in fields like computer-aided diagnosis [@lee_image_2015], pedestrian detection [@brunetti_computer_2018], and facial recognition [@emami_facial_2012], their application in reading data plots remains limited. While some studies have explored the use of computer vision models for tasks such as reading recurrence plots for time series regression [@ojeda_multivariate_2020], time series classification [@chu_automatic_2019; @hailesilassie_financial_2019; @hatami_classification_2018; @zhang_encoding_2020], anomaly detection [@chen_convolutional_2020], and pairwise causality analysis [@singh_deep_2017], the application of reading residual plots with computer vision models represents a relatively new field of study.

@@

### A plot is worth a thousand tests: assessing residual diagnostics with the lineup protocol

In order to collect data of human performance on reading residual plot of linear regression model and examine the potential applicability of integrating visual inference techniques into regression diagnostics, a pilot study was conducted in the first year with the participation of 64 individuals, which was followed by a formal study involving 443 participants in the subsequent year. The participants were presented with lineups consisting of 20 plots, where a residual plot was embedded along with 19 null plots, drawn with data simulated using the residual rotation technique. The study considered two primary forms of residual departures in multiple linear regression, namely non-linearity and heteroskedasticity. To enrich the visual features, different fitted value distributions, including normal, uniform, log-normal, and discrete, were also incorporated.

The study revealed that appropriate conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions compared to visual tests evaluated by humans, leading to excessive rejections even when downstream analysis and outcomes would not be significantly impacted by the small departures from a good fit. Specifically, conventional tests conclude that the model fit has issues nearly twice as frequently as humans would, and they frequently reject departures in the form of non-linearity and heteroskedasticity that are not visible to humans.

These findings emphasize the crucial role of graphical diagnostics and support the integration of visual inference into regression diagnostics. Moreover, they provide a compelling rationale for the development of an automated visual inference system to evaluate lineups of residual plots. The comprehensive details of the research methodology, along with other intriguing findings can be found in @li2023plot.

### Automated assessment of residual plots with computer vision models

In the third year of the project, a series of endeavors were undertaken to develop a computer vision model for assessing residual plots. To properly train such a model, a distance measure based on Kullback-Leibler divergence was introduced. This distance aimed to quantify the disparity between the residual distribution of a fitted linear regression model and the reference residual distribution assumed under correct model specification. It effectively captured the magnitude of model violations in a misspecified model.

Subsequently, the computer vision model was devised to approximate this distance by taking a residual plot of the fitted model as input. The resulting approximated distance predicted by the model could be utilized to construct a Model Violation Index for quantifying various model violations. Additionally, a formal statistical testing procedure was developed based on the approximated distance. This procedure produced critical values by evaluating a large number of null plots generated from the fitted model and estimating the sample quantiles of the distribution of the approximated distance. Furthermore, a bootstrapping technique was developed to measure how frequently the fitted model was considered misspecified if data were repeatedly obtained from the same data generating process.

The proposed computer vision model was trained exclusively on synthetic data, encompassing residual plots featuring non-linearity, heteroskedasticity, and non-normality issues. The trained model exhibited strong performance on both training and test sets. Statistical tests relying on the approximated distance predicted by the computer vision model showcased lower sensitivity compared to conventional tests but higher sensitivity compared to visual tests conducted by humans. The approximated distance generally aligned with the strength of the visual signal perceived by humans.

Through various examples, the effectiveness of the proposed method were illustrated across diverse scenarios, highlighting the similarity between visual tests and tests based on approximated distance. Furthermore, cases where conventional tests might fail, whereas visual tests and tests based on approximated distance remain effective were underscored.

This method holds significant value as it alleviates a portion of analysts' workload associated with assessing residual plots. While plotting residuals and evaluating them by analysts are still recommend for their invaluable insights, the trained computer vision model provided in this study can serve as a tool for automating the diagnostic process or for supplementary purposes as needed. The comprehensive details of the distance definition, model training process, along with other results can be found in the attached paper.

@@

2. Chapter Two presents a comparative study between visual testing and conventional testing in the context of regression diagnostics. The corresponding paper is already accepted by the Journal of Computational and Graphical Statistics.

3. Chapter Three discusses the development of an automatic visual inference system for evaluating lineups of residual plots. The computer vision models used in this chapter are primarily trained on simulated data from linear models that violate classical assumptions such as linearity and homoscedasticity. The chapter concludes with an evaluation of the performance of the system, comparing the power and characteristics of human subject-assessed visual tests with computer-assessed visual tests. This chapter will be identical to the attached paper.

4. Chapter Four provides detailed description of the open-source software developed during the course of the research. This includes the R package `bandicoot` for code maintenance with object-oriented programming, `visage` for running visual inference experiments on linear regression models, and `autovi` for assessing residual plots with trained computer vision models. Additionally, an online application of the automatic visual inference system is introduced.

@@

The timetable of my PhD study from Mar 2024 to August 2024 is provided in Table \ref{tab:timetable}.

```{r}
data.frame(Date = c("Mar",
                    "April", 
                    "May",
                    "July",
                    "",
                    "Aug"), 
           Event = c("Finish and submit the second paper",
                     "Develop a web interface for the automatic visual inference system",
                     "Clean and publish R packages to CRAN",
                     "Finish the third paper",
                     "Attend the useR! conference",
                     "Submit thesis")) %>%
  kableExtra::kable(booktabs = TRUE,
                    label = "timetable",
                    caption = "Timetable till thesis submission") %>%
  kableExtra::kable_styling(latex_options = "scale_down") 
```

### 2024-02-19 Milestone_3/report/report.tex
update milestone 3

\makeatletter
\def\organization#1{\gdef\@organization{#1}}
\def\telephone#1{\gdef\@telephone{#1}}
\def\email#1{\gdef\@email{#1}}
\makeatother
  \organization{Progress review}

@@

keywords:
AI, data visualization, residual plot, visual inference, hypothesis testing, computer vision

@@

The reliance of human assessment is a fundamental aspect of visual tests, but it may restrict its widespread usage. The lineup protocol is unsuitable for large-scale applications, due to its high labor costs and time requirements. Moreover, it presents significant usability issues for individuals with visual impairments, resulting in reduced accessibility.

Modern computer vision models offer a promising solution to this challenge. As a subfield of AI, computer vision with its modern deep learning architectures has successfully resolved numerous critical problems in automation. The development of the convolutional neural network (CNN) by \textcite{fukushima_neocognitron_1982} was inspired by the vision processing in living organisms. The modern computer vision model typically utilizes deep neural networks with convolutional layers, which leverage the hierarchical pattern in data and provide regularized versions of fully-connected layers. This approach downscales and transforms images by summarizing information in a small space. Numerous studies have shown that it can effectively tackle vision tasks, such as image recognition \autocite{rawat_deep_2017}, computer-aided diagnosis \autocite{lee_image_2015}, pedestrian detection \autocite{brunetti_computer_2018}, and facial recognition \autocite{emami_facial_2012}.

Utilizing computer vision models in reading data plot is not a common choice. Nevertheless, certain fields have adopted this idea by applying computer vision models in reading recurrence plots for time series regression \autocite{ojeda_multivariate_2020}, time series classification \autocite{chu_automatic_2019,hailesilassie_financial_2019,hatami_classification_2018,zhang_encoding_2020}, anomaly detection \autocite{chen_convolutional_2020}, and pairwise causality analysis \autocite{singh_deep_2017}. However, the assessment of lineups with computer vision models is a relatively novel area of study.

@@

In order to examine the potential applicability of integrating visual inference techniques into regression diagnostics, a pilot study was conducted in the first year with the participation of 64 individuals, which was followed by a formal study involving 443 participants in the subsequent year. The participants were presented with lineups consisting of 20 plots, where a residual plot was embedded along with 19 null plots, drawn with data simulated using the residual rotation technique. The study considered two primary forms of residual departures in multiple linear regression, namely non-linearity and heteroskedasticity. To enrich the visual features, different fitted value distributions, including normal, uniform, lognormal, and discrete, were also incorporated.

As shown in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower}, the study revealed that appropriate conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions compared to visual tests evaluated by humans, leading to excessive rejections even when downstream analysis and outcomes would not be significantly impacted by the small departures from a good fit. Specifically, conventional tests conclude that the model fit has issues nearly twice as frequently as humans would, and they frequently reject departures in the form of non-linearity and heteroskedasticity that are not visible to humans. For instance, based on the row of scatter plots at the bottom of Figure \ref{fig:polypower}, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5. The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25.

\begin{figure}
\centering
\includegraphics{polypower-1.png}
\caption{Comparison of power between different tests for detecting non-linearity. The power curves are estimated using logistic regression, and the horizontal lines of dots represent non-reject and reject results from human observers for each lineup. The visual test has multiple power curves estimated from bootstrap samples. The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.\label{fig:polypower}}
\end{figure}

\begin{figure}
\centering
\includegraphics{heterpower-1.png}
\caption{Comparison of power between different tests for detecting heteroskedasticity. Main plot shows the power curves, with dots indicating non-reject and reject in visual testing of lineups. The multiple lines for the visual test arise from estimating the power on many bootstrap samples. The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.\label{fig:heterpower}}
\end{figure}

These findings emphasize the crucial role of graphical diagnostics and support the integration of visual inference into regression diagnostics. Moreover, they provide a compelling rationale for the development of an automated visual inference system to evaluate lineups of residual plots. The comprehensive details of the research methodology, along with other intriguing findings can be found in the attached paper.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Chapter One provides an overview of regression diagnostics, visual inference, and computer vision models. This chapter serves as an introductory section of the thesis and forms the basis for subsequent discussions.
\item
  Chapter Two presents a comparative study between visual testing and conventional testing in the context of regression diagnostics. This chapter will be identical to the attached paper.
\item
  Chapter Three discusses the development of an automatic visual inference system for evaluating lineups of residual plots. The computer vision models used in this chapter are primarily trained on simulated data from linear models that violate classical assumptions such as linearity and homoscedasticity. The chapter concludes with an evaluation of the performance of the system, comparing the power and characteristics of human subject-assessed visual tests with computer-assessed visual tests.
\item
  Chapter Four provides detailed description of the open-source software developed during the course of the research. This includes the R package \texttt{bandicoot} for code maintenance with object-oriented programming, \texttt{visage} for running visual inference experiments on linear regression models, and one or multiple R packages for employment of the automatic visual inference system. Additionally, an online application of the automatic visual inference system is introduced. This chapter may be splitted into multiple chapters to provide a more in-depth description of each tool.
\item
  Chapter Five offers a comprehensive discussion of the thesis, emphasizing the implications and contributions of the research. The chapter also addresses the limitations of the research and outlines potential future research directions.
\end{enumerate}

@@

The timetable of my PhD study from April 2023 to August 2024 is provided in Table \ref{tab:timetable}.

@@

\caption{\label{tab:timetable}Timetable till thesis submission}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{ll}
\toprule
Date & Event\\
\midrule
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{2023}}\\
\hspace{1em}April & Submit the abstract of the first paper to Australian Statistical Conference (ASC)\\
\hspace{1em}May & Submit the first paper\\
\hspace{1em}June & Submit a poster for the first paper to Institute of Electrical and Electronics Engineers (IEEE) Visualization Conference\\
\hspace{1em}July & Leave for a month\\
\hspace{1em}Sep & Finalize the computer vision model\\
\hspace{1em}Oct & Attend the IEEE Visualization Conference\\
\hspace{1em}Nov & Start the web interface development\\
\hspace{1em}Dec & Attend the ASC\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{2024}}\\
\hspace{1em}Mar & Submit the second paper\\
\hspace{1em}Aug & Submit thesis\\
\bottomrule
\end{tabular}}
\end{table}

>>>>

\makeatletter
\def\organization#1{\gdef\@organization{#1}}
\def\telephone#1{\gdef\@telephone{#1}}
\def\email#1{\gdef\@email{#1}}
\makeatother
  \organization{Pre-submission review}

@@

@@

The reliance of human assessment is a fundamental aspect of visual tests, but it may restrict its widespread usage. The lineup protocol is unsuitable for large-scale applications, due to its high labour costs and time requirements. Moreover, it presents significant usability issues for individuals with visual impairments, resulting in reduced accessibility.

Modern computer vision models are well-suited for addressing this challenge. They rely on deep neural networks with convolutional layers \autocite{fukushima_neocognitron_1982}. These layers leverage hierarchical patterns in data, downsizing and transforming images by summarizing information in a small space. Numerous studies have demonstrated the efficacy of convolutional layers in addressing various vision tasks, including image recognition \autocite{rawat_deep_2017}. Despite the widespread use of computer vision models in fields like computer-aided diagnosis \autocite{lee_image_2015}, pedestrian detection \autocite{brunetti_computer_2018}, and facial recognition \autocite{emami_facial_2012}, their application in reading data plots remains limited. While some studies have explored the use of computer vision models for tasks such as reading recurrence plots for time series regression \autocite{ojeda_multivariate_2020}, time series classification \autocite{chu_automatic_2019,hailesilassie_financial_2019,hatami_classification_2018,zhang_encoding_2020}, anomaly detection \autocite{chen_convolutional_2020}, and pairwise causality analysis \autocite{singh_deep_2017}, the application of reading residual plots with computer vision models represents a relatively new field of study.

@@

\hypertarget{a-plot-is-worth-a-thousand-tests-assessing-residual-diagnostics-with-the-lineup-protocol}{%
\subsubsection{A plot is worth a thousand tests: assessing residual diagnostics with the lineup protocol}\label{a-plot-is-worth-a-thousand-tests-assessing-residual-diagnostics-with-the-lineup-protocol}}

In order to collect data of human performance on reading residual plot of linear regression model and examine the potential applicability of integrating visual inference techniques into regression diagnostics, a pilot study was conducted in the first year with the participation of 64 individuals, which was followed by a formal study involving 443 participants in the subsequent year. The participants were presented with lineups consisting of 20 plots, where a residual plot was embedded along with 19 null plots, drawn with data simulated using the residual rotation technique. The study considered two primary forms of residual departures in multiple linear regression, namely non-linearity and heteroskedasticity. To enrich the visual features, different fitted value distributions, including normal, uniform, log-normal, and discrete, were also incorporated.

The study revealed that appropriate conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions compared to visual tests evaluated by humans, leading to excessive rejections even when downstream analysis and outcomes would not be significantly impacted by the small departures from a good fit. Specifically, conventional tests conclude that the model fit has issues nearly twice as frequently as humans would, and they frequently reject departures in the form of non-linearity and heteroskedasticity that are not visible to humans.

These findings emphasize the crucial role of graphical diagnostics and support the integration of visual inference into regression diagnostics. Moreover, they provide a compelling rationale for the development of an automated visual inference system to evaluate lineups of residual plots. The comprehensive details of the research methodology, along with other intriguing findings can be found in \textcite{li2023plot}.

\hypertarget{automated-assessment-of-residual-plots-with-computer-vision-models}{%
\subsubsection{Automated assessment of residual plots with computer vision models}\label{automated-assessment-of-residual-plots-with-computer-vision-models}}

In the third year of the project, a series of endeavors were undertaken to develop a computer vision model for assessing residual plots. To properly train such a model, a distance measure based on Kullback-Leibler divergence was introduced. This distance aimed to quantify the disparity between the residual distribution of a fitted linear regression model and the reference residual distribution assumed under correct model specification. It effectively captured the magnitude of model violations in a misspecified model.

Subsequently, the computer vision model was devised to approximate this distance by taking a residual plot of the fitted model as input. The resulting approximated distance predicted by the model could be utilized to construct a Model Violation Index for quantifying various model violations. Additionally, a formal statistical testing procedure was developed based on the approximated distance. This procedure produced critical values by evaluating a large number of null plots generated from the fitted model and estimating the sample quantiles of the distribution of the approximated distance. Furthermore, a bootstrapping technique was developed to measure how frequently the fitted model was considered misspecified if data were repeatedly obtained from the same data generating process.

The proposed computer vision model was trained exclusively on synthetic data, encompassing residual plots featuring non-linearity, heteroskedasticity, and non-normality issues. The trained model exhibited strong performance on both training and test sets. Statistical tests relying on the approximated distance predicted by the computer vision model showcased lower sensitivity compared to conventional tests but higher sensitivity compared to visual tests conducted by humans. The approximated distance generally aligned with the strength of the visual signal perceived by humans.

Through various examples, the effectiveness of the proposed method were illustrated across diverse scenarios, highlighting the similarity between visual tests and tests based on approximated distance. Furthermore, cases where conventional tests might fail, whereas visual tests and tests based on approximated distance remain effective were underscored.

This method holds significant value as it alleviates a portion of analysts' workload associated with assessing residual plots. While plotting residuals and evaluating them by analysts are still recommend for their invaluable insights, the trained computer vision model provided in this study can serve as a tool for automating the diagnostic process or for supplementary purposes as needed. The comprehensive details of the distance definition, model training process, along with other results can be found in the attached paper.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Chapter One provides an overview of regression diagnostics, visual inference, and computer vision models. This chapter serves as an introductory section of the thesis and forms the basis for subsequent discussions.
\item
  Chapter Two presents a comparative study between visual testing and conventional testing in the context of regression diagnostics. The corresponding paper is already accepted by the Journal of Computational and Graphical Statistics.
\item
  Chapter Three discusses the development of an automatic visual inference system for evaluating lineups of residual plots. The computer vision models used in this chapter are primarily trained on simulated data from linear models that violate classical assumptions such as linearity and homoscedasticity. The chapter concludes with an evaluation of the performance of the system, comparing the power and characteristics of human subject-assessed visual tests with computer-assessed visual tests. This chapter will be identical to the attached paper.
\item
  Chapter Four provides detailed description of the open-source software developed during the course of the research. This includes the R package \texttt{bandicoot} for code maintenance with object-oriented programming, \texttt{visage} for running visual inference experiments on linear regression models, and \texttt{autovi} for assessing residual plots with trained computer vision models. Additionally, an online application of the automatic visual inference system is introduced.
\item
  Chapter Five offers a comprehensive discussion of the thesis, emphasizing the implications and contributions of the research. The chapter also addresses the limitations of the research and outlines potential future research directions.
\end{enumerate}

@@

The timetable of my PhD study from Mar 2024 to August 2024 is provided in Table \ref{tab:timetable}.

@@

\caption{\label{tab:timetable}Timetable till thesis submission}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{ll}
\toprule
Date & Event\\
\midrule
Mar & Finish and submit the second paper\\
April & Develop a web interface for the automatic visual inference system\\
May & Clean and publish R packages to CRAN\\
July & Finish the third paper\\
 & Attend the useR! conference\\
\addlinespace
Aug & Submit thesis\\
\bottomrule
\end{tabular}}
\end{table}
